\textbf{Generalized additive models to analyze non-linear trends in biomedical longitudinal data using R: Beyond repeated measures ANOVA and Linear Mixed Models}

### 2022-01-20 Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/response_to_reviewers/Response_to_reviewers.Rmd

The essence of this is not “if the difference between [credible] intervals of the fitted smooths… is non-zero”. For a variety of reasons this is not how it works. Instead we compute the difference of the smooths and then compute a credible interval around this difference. Also, to interpret this correctly we really need to do a simultaneous interval as you are looking at all points over the range of the covariate. If we just use the usual interval we have to consider the under and over coverage of the Nychka across-the-function intervals. You should talk about this issue in this section if you don’t compute simultaneous intervals.
This section could also do with a little more explanation; {gratia} (and hence I) tens to consider differences of smooths to exclude the groups means. Here, if I read the paper and the code correctly, you included the group means. This works OK for simple models like this, but what if you had two or more parametric factor terms in the model? Then you would not (at least by default) have any way to isolate the group means for the by factor independently of the levels of the other factor. You would have to create differences conditional upon a level of the other factor. Users, however, tend to expect differences to include the group means, as you have chosen to present things here. Hence a note explaining the implications of this choice would be useful to forewarn the user.

### \textcolor{reviewersblue} {Reply to Reviewer's Comments}

We have made major changes to this section based on the reviewer's comments.

- We have removed the sentence "if the difference between the credible intervals of the fitted smooths... is non-zero", and now L491-L493 now read: "is that the difference between them can be computed, followed by the estimation of an empirical Bayesian simultaneous CI around this difference."

@@

### \textcolor{reviewersblue} {Reply to Reviewer's Comments}

We have changed the word from  "global" to "population".

### Reviewer's Comments

L283 I think it is better to say “…the mathematical space within which the true but unknown $f(time_t)$ is thought to exist.”

@@

### Reviewer's Comments

L356 use straight quotes in code, not curly ones. And speficy the `family` here?

### \textcolor{reviewersblue} {Reply to Reviewer's Comments}

We use the code notation provided by the $LaTeX$ package `listings`, and have updated the quotes to be straight in all code chunks. Regarding `family`, we believe that it is not necessary to include it here, but we indicate in L460-L461 what the argument exists and that other distributions can be used.

### 2022-01-20 Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/response_to_reviewers/Response_to_reviewers.tex

The essence of this is not ``if the difference between {[}credible{]} intervals of the fitted smooths\ldots{} is non-zero''. For a variety of reasons this is not how it works. Instead we compute the difference of the smooths and then compute a credible interval around this difference. Also, to interpret this correctly we really need to do a simultaneous interval as you are looking at all points over the range of the covariate. If we just use the usual interval we have to consider the under and over coverage of the Nychka across-the-function intervals. You should talk about this issue in this section if you don't compute simultaneous intervals.
This section could also do with a little more explanation; \{gratia\} (and hence I) tens to consider differences of smooths to exclude the groups means. Here, if I read the paper and the code correctly, you included the group means. This works OK for simple models like this, but what if you had two or more parametric factor terms in the model? Then you would not (at least by default) have any way to isolate the group means for the by factor independently of the levels of the other factor. You would have to create differences conditional upon a level of the other factor. Users, however, tend to expect differences to include the group means, as you have chosen to present things here. Hence a note explaining the implications of this choice would be useful to forewarn the user.

\hypertarget{section-9}{%
\subsubsection{\texorpdfstring{\textcolor{reviewersblue} {Reply to Reviewer's Comments}}{}}\label{section-9}}

We have made major changes to this section based on the reviewer's comments.

\begin{itemize}
\item
  We have removed the sentence ``if the difference between the credible intervals of the fitted smooths\ldots{} is non-zero'', and now L491-L493 now read: ``is that the difference between them can be computed, followed by the estimation of an empirical Bayesian simultaneous CI around this difference.''
\item
  L497-L498 The sentence now reads ``unlikely to be distinguishable from zero''.
\item
  L507-513 now indicate that we are computing a simultaneous CI around the difference, and that we \emph{included} the group means in our pairwise comparison. We also indicate to the reader the fact that if multiple parametric terms exist in the model the inclusion of the means can become problematic.
\end{itemize}

@@

\hypertarget{section-22}{%
\subsubsection{\texorpdfstring{\textcolor{reviewersblue} {Reply to Reviewer's Comments}}{}}\label{section-22}}

We have changed the word from ``global'' to ``population''.

\hypertarget{reviewers-comments-22}{%
\subsubsection{Reviewer's Comments}\label{reviewers-comments-22}}

L283 I think it is better to say ``\ldots the mathematical space within which the true but unknown \(f(time_t)\) is thought to exist.''

@@

\hypertarget{reviewers-comments-26}{%
\subsubsection{Reviewer's Comments}\label{reviewers-comments-26}}

L356 use straight quotes in code, not curly ones. And speficy the \passthrough{\lstinline!family!} here?

\hypertarget{section-27}{%
\subsubsection{\texorpdfstring{\textcolor{reviewersblue} {Reply to Reviewer's Comments}}{}}\label{section-27}}

We use the code notation provided by the \(LaTeX\) package \passthrough{\lstinline!listings!}, and have updated the quotes to be straight in all code chunks. Regarding \passthrough{\lstinline!family!}, we believe that it is not necessary to include it here, but we indicate in L460-L461 what the argument exists and that other distributions can be used.

### 2022-01-20 Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/response_to_reviewers/Response_to_reviewers.tex

We would like to thank the reviewers for the careful and thorough analysis of this manuscript to \emph{Statistics in Medicine} and for the thoughtful comments and constructive suggestions, which helped improve the quality of this manuscript. We carefully considered the reviewers comments and in this document, explain how we revised the manuscript based on those comments and suggestions.

\hypertarget{general-comments-to-the-editor}{%
\section{General Comments to the Editor}\label{general-comments-to-the-editor}}

Dr.~Platt,

The authors thank you for your determination that our manuscript may be suitable for resubmission in \emph{Statistics in Medicine} after addressing the reviewer's comments. To this end, we have addressed all critiques. We hope these revisions, submitted on January 20th, 2022, improve the manuscript so it is deemed worthy of publication in \emph{Statistics in Medicine}. Following are our detailed responses to reviewer comments.
